\textcolor{red}{Simple} connectome inference using partial correlation from imaging calcium signal

### 2014-06-16 article.tex
wording in filtering

The directivity in the network introduce causal relations through neurons. A directed edge, i.e. $a_{i,j} = 1$ and $a_{j,i} = 0$, means that neuron $i$ causes $j$. The causal mechanism implies that when neuron $i$ is active, also denoted in this paper as a firing event or a burst, neuron $j$ will be consequently active aftewards. The reciprocal phenomenon is only possible when the edge is bidirectionnal, i.e. $a_{i,j} = 1 = a_{j,i} = 1$.

@@

The raw signal is the calcium fluorescence imaging representing
neuron activity. Considering the imaging process, we have to cope with a slower
sampling rate than the neuron firing frequency and the slow decay of fluorescent
calcium indicator after a neuron activity burst.

@@

As it can be seen on \textit{fig 1}, the raw signal is very noisy and
a low-pass filter is therefore applied to get a cleaner version of the signal where only
true variations are kept, i.e. without noise variations. However, the choice of a low-pass
filter is not unique. In the rest of the paper, we will focus on two types, the
symmetrical median filter, i.e., $x^t_i \leftarrow x^{t-1}_i + x^{t}_i + x^{t+1}_i$,
and an asymmetrical weighted median filter, i.e.,
$x^t_i \leftarrow 0.4 x^{t-3}_i + 0.6 x^{t-2}_i + 0.8 x^{t-1}_i + x_{t}^i$.

@@

The small decay for the fluorescent calcium
indicator implies a signal variation during a long interval and therefore, in
order to only keep rising and strong transitions, we apply a hard-tresholing filter:
$x^{t,d}_i = X^{t,d}_i \mathbb{1}(x^{t,d}_i \geq \tau)$ where $\tau$ is a given threshold 
As a result, we get a binary-like signal with the true activity when a burst
happens and with a true zero when the neuron is inactive.

@@

Then we apply on the entire signal a non-linear filter based on the
current activity of the systems in order to favor the association measure inversely
proportional to the number of neurons firing together:
$ X_{t,w}^i  = (x^{t,d}_i + 1 )^{1 + \frac{1}{\sum_{j} X_{t,d}^j}} $.
When too much neurons are firing together, i.e. in case of a network burst, it is almost
impossible to retrieve direct relationships: all neurons seem to be directly
connected to each other. By favoring the association measure when the activity is low,
we expect to retrieve a pair of neurons that are really directly connected together.
In concrete terms, as it can be verified on \textit{fig 1}, peak intensity may change in function
of the activity of all neurons but the number and the locations of peaks is not modified.

>>>>

The directivity in the network introduce causal relations through neurons. A directed edge, i.e. $a_{i,j} = 1$ and $a_{j,i} = 0$, means that neuron $i$ causes $j$. The causal mechanism implies that when neuron $i$ is active, also denoted in this paper as a firing event or a burst, neuron $j$ will be afterwards active consequently. The reciprocal phenomenon is only possible when the edge is bidirectionnal, i.e. $a_{i,j} = a_{j,i} = 1$.

@@

The raw signal is the calcium fluorescence imaging representing neuron
activity via a capture imaging process. Observing the fluorescent calcium as
an indicator of neuron activity delivers a 2-D view of neuron network at each
capture which only gives a network screenshot at given time points. It implies
obvious issues. Some neurons may be invisible because they are not observed or
they are confounded with another one because of the loss of the third
dimension. The capture sample rate is also a constraint that could make
difficult to detect in which order the neurons fire.  Moreover the activity
indicator is highly reactive to a burst but takes some time to recover its non
active state.  Finally, the difficulty to distinguish direct relationships when there is a network burst, defined as a highly synchronous activity [Stetter and in Stetter [20,46,47]], is increased by the slow sampling rate.

Considering the described imaging process, we try in this
section to cope with some of these issues and, in particular, the slower
sampling rate than the neuron firing frequency and the slow decay of
fluorescent calcium indicator after a neuron activity burst.

@@

As it can be seen on \textit{fig 1}, the raw signal is very noisy due to light scattering artifacts that ordinarily affects the quality of the recording [in Stetter, [6]]. Therefore, a low-pass filter is first applied to get a cleaner version of the signal where only
true variations are kept, i.e. without noise variations. However, the choice of a low-pass
filter is not unique. In the rest of the paper, we will use two particular filters giving similar results, the
symmetrical median filter, i.e., $x^t_i \leftarrow x^{t-1}_i + x^{t}_i + x^{t+1}_i$,
and an asymmetrical weighted median filter, i.e.,
$x^t_i \leftarrow 0.4 x^{t-3}_i + 0.6 x^{t-2}_i + 0.8 x^{t-1}_i + x_{t}^i$.

@@

The small decay for the fluorescent calcium indicator implies a signal
variation during a long time interval. Firing event are represented by 
strong variations in the neuron activity. Therefore we apply a hard-tresholding 
filter $x^{t,d}_i = x^{t,d}_i \mathbb{1}(x^{t,d}_i \geq \tau)$ where the threshold $\tau$ should be taken positive to keep only rising edges and non close to zeros in order to only consider strong variations.

As a result, we get a binary-like signal with the true activity when a burst
happens and with a true zero when the neuron is inactive.

@@

Then we apply on the entire signal a non-linear filter based on the
current activity of the systems in order to favor the association measure inversely
proportional to the number of neurons firing together:
$ x^{t,w}_i  = (x^{t,d}_i + 1 )^{1 + \frac{1}{\sum_{j} X_{t,d}^j}} $.
When too much neurons are firing together, i.e. in case of a network burst, it is almost
impossible to retrieve direct relationships: all neurons seem to be directly
connected to each other. By favoring the association measure when the activity is low,
we expect to retrieve a pair of neurons that are really directly connected together.
In concrete terms, as it can be verified on \textit{fig 1}, peak intensity may change in function
of the activity of all neurons but the number and the locations of peaks is not modified.
